Theoretical prediction of disrupted Min oscillation in flattened \emph{Escherichia coli}

### 2015-06-21 paper/response-to-referees.txt

Dear Editor,

We thank the referees for their helpful comments.  We have made a
number of revisions to our paper in response to their suggestions,
which we enumerate below, along with our response to each comment and
suggestion of the referees.  Each referee had several major concerns
and a few minor suggestions, which we will address in the order that
they were given, beginning with Referee #1.

Referee #1.

1. The largest concern of Referee #1 was our choice of Huang et al's
model for our simulations.  The referee acknowledges that we can stay
with this model, but feels that our justifaction that this model is
inadequate.  The referee objects that the fact that this model has
been widely is an inappropriate justification.  While we agree that
this alone is not a sufficient reason for studying this model, we do
believe that it is relevant: this is not some obscure older model that
we are testing, but rather the basis of all more recent models.  We
believe that the limitations of this model are of interest, and have
added some discussion to the end of the introduction section
explaining this.

@@

2. As a secondary point, the referee points out we cannot conclude
that some other deterministic model could not demonstrate chaotic
behavior that explains the experimental results of Mannik et al.  We
agree that this cannot be ruled out, and have added mention of this to
our "Comparison of Experimental and Stadium Shapes", and have made
even more tentative our suggestion in the summary that other similar
deterministic models may also prove to be non-chaotic.

3. The referee when looking at Figure 4 observes that the stadium
shapes exhibit osciallation that is largely bipolar, and expresses
doubt in our claim that the oscillations in stadium shapes are
irregular.  In particular, the referee points out that if we average
over many periods, we would find that the oscillations are indeed
bipolar.  We agree with this statement, but not with the conclusion.
Averaging the locations of the maxima would indeed miss the point of
the comparison with experiment, which is in examining the sequence of
locations of maxima.

We have edited the "comparison of experimental and stadium shapes"
section to moderate our claims regarding the spatial irregularity of the
experimental and stadium shapes as predicted by the stochastic model.

<Jeff will figure out what other places need to be changed to reflect
this reduced claim. EDIT ELSEHWERE AND EDIT HERE!>

@@

6. The referee points out that two previous works have also
demonstrated that stochastic simulations are needed in order to
explain WHAT?.  WE HAVE EDITED OUR DISCUSSION TO ILLUSTRATE THIS.

  <David does this.>

Referee #3.

7. The referee #3 also expressed doubt about our conclusion with
regard to the effect of irregular shape which we base on our stadium
models.  Unlike the previous referee, the criticism is based not on
the spatial output (Figure 4), but rather on the reported coherence
times.  We agree with the referee that our concluding sentence
immediately prior to the conclusion section was unjustified, and have
edited this sentence to reflect the referee's objections.

<Jeff will check if we talk about decoherence times later>

@@

8. The referee also questioned how the cell was divided in order to
perform the coherence time calculations, pointing out that a smaller
region could lead to a shorter coherence time.  EXPLAIN HOW WE DIVIDE
THE CELLS UP IN THIRDS BY VOLUME? IT'S AWKWARD THAT WE DON'T DIVIDE
ALONG LONG AXIS.  JEFFSAYS{Bad news here.  It looks like I did this
dividing up the cells by eye.  Originally the slices were used to
created the box plots, which was only a way to visualize the data in
smaller time domains, so there's wasn't really any disadvantage to
doing it roughly like this (we even had the visually appealing little
colored diagrams on the sides of the box plots).  Not sure what to say
abou this in text.}

  <Jeff needs to find out how much volume in each section! Then write
  text explaining this...>

@@

9. The referee points out that the "starry night" patterns are
inconsistent with experiments.  We agree with the referee that
membrane diffusion will alleviate this problem.  WE HAVE ADDED A
CLARIFICATION TO THIS EFFECT WHERE?

  <David>

Minor points:

10. As the referee requested, we have clarified the caption of Figure
1, to include the fact that these reactions are for a Min system
model that is used in this work.

@@

11. The referee suggests that we should write $\mu m^2$ over $\mu^2$
for square microns.  We have made this change.  FIXME DO SO.

  <David>

12. The referee suggested a title is needed in the caption of Figure
number 4.  We have edited the caption to begin with an appropriate
title-like sentence fragment, rather than "We display here...".

### 2015-06-21 paper/plos.tex
edits from talking with Jeff

%% biology~\cite{lawson2013spatial, robb2014stochastic,
%%   oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
In this paper we use Huang's original 2003
model~\cite{huang2003dynamic}, which has been widely
used~\cite{hattne2005stochastic,huang2004min,kerr2006division,
  fange2006noise} in both deterministic and stochastic variants, to
study flattened cells of $0.4\micron$ thickness that are similar to
those observed by M\"annik \emph{et al.}~\cite{mannik2009bacterial}.
\fixme{This allows us to test the performance of this important model
  in more extreme situations than has been done previously, furthering
  considerably our understanding of how and to what extent it breaks
  down. It will be valuable in the future to compare these results
  against similar tests of more recent models, in order to better
  understand how they gain advantage.}

\begin{figure}
  \includegraphics[width=8.7cm]{reactions}
  \caption{Reactions included in the model of Huang \emph{et
      al.}~\cite{huang2003dynamic}.}\label{fig:reactions}
\end{figure}

@@

The predictions of the stochastic model are roughly similar to the
experimental results in terms of irregularity of the locations of
maxima, although there are locations on the edges of the cells where
experiment shows maxima occurring that we never see in our
simulations, suggesting that the model does not precisely reflect the
experiment.  We also note that the stadium shapes in
Fig.~\ref{randst-plot-ave} appear to have maxima location irregularity
\fixme{within the time domain of a small number of periods, as
  compared to the more long term coherence discussed below,} that is
qualitatively similar to our irregular shapes and to experiment.

@@

We see that both in the shape \emph{B} and stadium \emph{B}, the
deterministic model is perfectly periodic.  Both the deterministic and
stochastic model correlations have the same number of proteins in
their cells, and have been normalized with the same normalization
factor, so that they can be directly compared against each other.  The
stochastic simulation of shape \emph{B} has a coherence time of 5.3
periods, while the corresponding simulation of stadium \emph{B} has a
coherence time of 8.6 periods.
%
The correlation function for stadium \emph{A} is similar to that of
stadium~\emph{B}, with similar coherence time (13 periods, not shown),
while shape~\emph{A} also exhibits a similar coherence time (5.5
periods).  This confirms that the stadium shapes do indeed oscillate
somewhat more regularly than the irregular shapes, but both are
comparable to the wild-type pill shape.
%
From the stochastic correlation data, we conclude that irregularity of
shape has \fixme{a distinct yet moderate} affect on the temporal regularity of
the oscillatory behavior, with even the very distorted shape~\emph{B}
exhibiting only a factor of two greater decoherence.

@@

We find that flattened but regular and symmetric cells exhibit MinD
oscillations that are qualitatively similar to the oscillations
observed experimentally (and in our simulation) in irregular and
asymmetric cells.  This demonstrates that asymmetry is not required in
order to induce spatially irregular MinD oscillations.  In addition,
the temporal regularity of oscillations is only moderately affected by
irregularity in the shape of flattened cells.  Thus we conclude that
it is \fixme{primarily} the flattening of the cells, and the
accompanying expansion in the transverse direction, \fixme{deleted
  some text} that causes the disruption of regular bipolar oscillation
which is observed by M\"annik \emph{et
  al.}~\cite{mannik2012robustness}.

>>>>

%% biology~\cite{lawson2013spatial, robb2014stochastic,
%%   oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
In this paper we use Huang's original 2003
model~\cite{huang2003dynamic}, which has been widely
used~\cite{hattne2005stochastic,huang2004min,kerr2006division,
  fange2006noise} in both deterministic and stochastic variants, to
study flattened cells of $0.4\micron$ thickness that are similar to
those observed by M\"annik \emph{et al.}~\cite{mannik2009bacterial}.
This allows us to test the performance of this important model in more
extreme situations than has been done previously, furthering our
understanding of how and to what extent it breaks down.

\begin{figure}
  \includegraphics[width=8.7cm]{reactions}
  \caption{Reactions included in the Min system model of Huang
    \emph{et al.}, which is used in this
    work~\cite{huang2003dynamic}.}\label{fig:reactions}
\end{figure}

@@

The predictions of the stochastic model show some similarity to the
experimental results in terms of irregularity of the locations of
maxima, but the stochastic model displays significantly less
variation in the direction of oscillations than experiment, suggesting
inadequacies in the model.  We also note that the stadium shapes in
Fig.~\ref{randst-plot-ave} appear to have maxima location irregularity
that is qualitatively similar to our irregular shapes: in both
stochastic simulations the overall oscillation is largely bipolar, but
with irregularity in the locations of the maxima that form at each
end.

@@

We see that both in the shape \emph{B} and stadium \emph{B}, the
deterministic model is perfectly periodic.  Both the deterministic and
stochastic model correlations have the same number of proteins in
their cells, and have been normalized with the same normalization
factor, so that they can be directly compared against each other.  The
stochastic simulation of shape \emph{B} has a coherence time of 5.3
periods, while the corresponding simulation of stadium \emph{B} has a
coherence time of 8.6 periods.
%
The correlation function for stadium \emph{A} is similar to that of
stadium~\emph{B}, with similar coherence time (13 periods, not shown),
while shape~\emph{A} also exhibits a similar coherence time (5.5
periods).  This confirms that the stadium shapes do indeed oscillate
somewhat more regularly than the irregular shapes, but both are
comparable to the wild-type pill shape.
%
From the stochastic correlation data, we conclude that irregularity of
shape leads to a decrease in temporal regularity of the oscillatory
behavior, although even the very distorted shape~\emph{B} exhibiting
only a factor of two greater decoherence.

@@

We find that flattened but regular and symmetric cells exhibit MinD
oscillations that are qualitatively similar to the oscillations
observed experimentally (and in our simulation) in irregular and
asymmetric cells.  This demonstrates that asymmetry is not required in
order to induce spatially irregular MinD oscillations.  In addition,
the temporal regularity of oscillations is only moderately affected by
irregularity in the shape of flattened cells.  Thus we conclude that
it may be the flattening of the cells, and the accompanying
expansion in the transverse direction, that causes the disruption of
regular bipolar oscillation which is observed by M\"annik \emph{et
  al.}~\cite{mannik2012robustness}.
